---
title: "Import MedStar Compliance Data"
date: "Created: 2016-09-09 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    css: custom-css.css
---

# Table of contents

* [Overview](#overview)   
* [Load packages and data](#load)   
* [Import raw data](#import)   
* [Save DETECT pilot response id's](#detect-ids)   
* [Save compliance data](#save)     


# Overview

In this file we import the data sent to us by MedStar that includes every report to APS that their legal compliance department was aware of in 2015. Specifically, the input data set is called:

* APS Report Data_UNTHSC_IRB.xlsx

We also save a list of the response id numbers that correspond to 911 responses where a report was made to APS during the DETECT pilot phase and MedStar compliance was made aware of the report.

Finally, we create a deidentified version of the MedStar compliance data.


# Load packages and data {#load}

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(readxl)
library(bfuncs)
```










# Import raw data {#import}

```{r}
col_info <- c( 
# "col_name",     "col_type"
  "agency",       "text",
  "confim_num",   "text",
  "created",      "date",
  "reported",     "date",
  "electronic",   "text",
  "phone",        "text",
  "response_num", "text")

# Import data
medstar_compliance <- read_excel(
  path = "/Volumes/sph_research/Detect/APS Report Data_UNTHSC_IRB.xlsx",
  sheet = "owssvr",
  col_names = col_info[seq(1, length(col_info), 2)],
  col_types = col_info[seq(2, length(col_info), 2)],
  skip = 1) %>% 
  arrange(reported)
```

```{r}
about_data(medstar_compliance) # 137 observations and 7 variables in the data
```


## Remove all the CPS cases

```{r}
medstar_compliance <- filter(medstar_compliance, agency == "APS")
about_data(medstar_compliance) # 93 observations and 7 variables in the data 
```


## Remove 2016 cases

```{r}
medstar_compliance <- filter(medstar_compliance, reported <= as.POSIXct("2016-01-01"))
about_data(medstar_compliance) # 49 observations and 7 variables in the data
```


## Identify DETECT pilot period

```{r}
start <- as.POSIXct("2015-09-17")
end   <- as.POSIXct("2015-10-26")

medstar_compliance <- medstar_compliance %>%
  mutate(
    detect_pilot = (reported >= start & reported <= end)
  )
```

```{r}
about_data(medstar_compliance) # 49 observations and 8 variables
```

At this point, the medstar_compliance data contains 49 observations and 8 variables. All remaining reports were made to APS during the 2015 calendar year.


[top](#top)










# Save DETECT pilot response id's {#detect-ids}

These are the response id numbers that correspond to 911 responses where a report was made to APS during the DETECT pilot phase and MedStar compliance was made aware of the report.

```{r response_id}
response_ids <- medstar_compliance %>% 
  filter(reported >= start & reported <= end) %>% 
  select(response_num)

nrow(response_ids) # 14
```

There were 14 response id's with a report date between 2015-09-17 and 2015-10-26 (inclusive) in the data from MedStar compliance.


## Save response id's

We will look for these response id's in the DETECT screening data as well.

```{r}
feather::write_feather(response_ids, "/Volumes/sph_research/Detect/response_ids.feather")
```

[top](#top)










# Save compliance data {#save}

For analysis

```{r}
feather::write_feather(
  medstar_compliance, 
  "/Volumes/sph_research/Detect/medstar_compliance.feather"
)
```

[top](#top)


&nbsp;

-------------------------------------------------------------------------------

```{r echo=FALSE}
sessionInfo()
```
