---
title: "Preprocess merged data for analysis"
date: "Created: 2018-10-27 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    css: custom-css.css
---

# Table of contents

* [Overview](#overview) 
* [Load packages and data](#load)   
* [Variable management](#var-manage)    
* [Save preprocessed data](#save)    

# Overview

In this file we take the MedStar DETECT and APS investigations data that we previously merged together and do a little bit of preprocessing for later analysis.


# Load packages and data {#load}

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(bfuncs)
```

```{r}
detect_5wk_1 <- feather::read_feather("../data/detect_5wk_1.feather")
```

```{r}
about_data(detect_5wk_1) # 98,968 observations and 52 variables
```










# Variable management {#var-manage}

Because it is possible for there to be more than one allegation associated with each APS investigation, some investigations have more than one row in the data. It is also possible to have more than one instance of the DETECT screening tool used (incident pcr number) per APS investigation (case number).

We want a single outcome per investigation/pcr combination in order to create a confusion matrix for each screening item.

To do that, we will create a new variable called "any_valid" that is equal to 1 if the disposition of _any_ allegation was valid, and 0 if the disposition of _all_ allegations were invalid. 

## Create any_valid

Create new outcome variable (any_valid) that is equal to 1 when any type of abuse is validated for a given response pair (DETECT screen / APS investigation match). For response pairs with multiple dispositions:   
  + If even one disposition is valid, then any valid equals 1   
  + If all dispositions are invalid, then any valid equals 0   
  + If all dispositions are other, then any valid equals NA
  + If one disposition is other, and the rest are invalid, then any valid equals NA
  
```{r}
detect_5wk_2 <- detect_5wk_1 %>%
  group_by(id) %>% 
  mutate(
    any_valid = case_when(
      any(disposition == "Valid")   ~ 1L,
      all(disposition == "Invalid") ~ 0L,
      TRUE                          ~ NA_integer_
    ) 
  ) %>% 
  ungroup()
```

## Data check

```{r}
detect_5wk_2 %>% 
  select(id, incident_pcr_number, case_num, disposition, any_valid) %>% 
  distinct()
```

```{r}
about_data(detect_5wk_2) # 98,968 observations and 53 variables
```


## Count number of screening items completed 

When the DETECT tool was used, it often wasn't the case that all screening items were used. Here, we create a variable that contains the number of screening items completed at each pcr, if any. For this purpose, "Not applicable or No Caregiver" and "Don’t Know" count as completed.

```{r}
detect_5wk_2 <- detect_5wk_2 %>% 
  mutate_at(
    .vars = vars(unusual_odor36:adls61),
    .funs  = funs(answered = if_else(!is.na(.), 1, 0))
  ) %>% 
  mutate(
    answered_count = select(., ends_with("answered")) %>% rowSums()
  ) %>% 
  select(-ends_with("answered"))
```

```{r}
about_data(detect_5wk_2) # 98,968 observations and 54 variables
```


## Recode DETECT responses to NA

* We will recode "Not applicable or No Caregiver" to NA. The only variable that contained this response was no_talk51. This response was given 3 times.

* We will also recode "Don’t Know" to NA.

```{r}
detect_5wk_2 <- detect_5wk_2 %>% 
  map_at(
    .x  = .,
    .at = select(detect_5wk_2, unusual_odor36:adls61) %>% names(),
    .f  = fct_recode, NULL = "Not applicable or No Caregiver"
  ) %>% 
  map_at(
    .x  = .,
    .at = select(detect_5wk_2, unusual_odor36:adls61) %>% names(),
    .f  = fct_recode, NULL = "Don’t Know"
  ) %>% 
  as_tibble()
```

```{r}
about_data(detect_5wk_2) # 98,968 observations and 54 variables
```


## Create any positive variable

Create a variable that is equal to 1 if any DETECT item was positive and equal to 0 if all screening items used were negative.

```{r}
detect_5wk_2 <- bind_cols(
  detect_5wk_2,
  detect_5wk_2 %>% 
    select(unusual_odor36:adls61) %>% 
    mutate_all(
      funs(
        yes = case_when(. == "Yes" ~ 1L, TRUE ~ 0L) # 1 if Yes, 0 if anything else - including NA
      )
    ) %>% 
    select(ends_with("yes")) %>% 
    mutate(
      yes_count = rowSums(.), # Count 1's within row across columns
      yes_any   = if_else(yes_count > 0, 1L, 0L),
      yes_any_f = factor(yes_any, labels = c("No", "Yes"))
    ) %>% 
    select(-ends_with("yes"))
)
```

```{r}
about_data(detect_5wk_2) # 98,968 observations and 57 variables
```

[top](#top)










# Save preprocessed data {#save}

This should be the version that will be posted on [NACJD]( https://www.icpsr.umich.edu/icpsrweb/content/NACJD/index.html)

```{r}
feather::write_feather(detect_5wk_2, "../data/detect_5wk.feather")
```


## Clean up

```{r}
rm(list = ls())
```

[top](#top)


&nbsp;

-------------------------------------------------------------------------------

```{r echo=FALSE}
sessionInfo()
```
