---
title: "Deidentify MedStar Compliance Data"
date: "Created: 2016-09-09 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    css: custom-css.css
---

# Table of contents

* [Overview](#overview)   
* [Load packages and data](#load)   
* [Remove days](#deidentify)   
* [Save deidentified data](#save)     


# Overview

Remove identifiers from the APS reporting data we received from MedStar's legal compliance department so that it can be uploaded to [NACJD]( https://www.icpsr.umich.edu/icpsrweb/content/NACJD/index.html)

As part of the process of protecting participant anonymity, we will remove the day from all dates.


# Load packages and data {#load}

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(bfuncs)
```

## Load data

Reporting data from MedStar's compliance department. Cleaned in data_medstar_compliance_01_import.Rmd

```{r}
medstar_compliance <- feather::read_feather("/Volumes/sph_research/Detect/medstar_compliance.feather")
```

```{r}
about_data(medstar_compliance) # 49 observations and 8 variables
```


# Remove days {#deidentify}

Remove identifiers from the APS reporting data we received from MedStar's legal compliance department so that it can be uploaded to [NACJD]( https://www.icpsr.umich.edu/icpsrweb/content/NACJD/index.html)

As part of the process of protecting participant anonymity, we will remove the day from all dates.

```{r}
medstar_compliance_deidentified <- medstar_compliance %>%
  select(created, reported, electronic, phone, detect_pilot) %>% 
  mutate(
    created  = format(created, "%Y-%m"),
    reported = format(reported, "%Y-%m")
  )
```

```{r}
about_data(medstar_compliance_deidentified) # 49 observations and 5 variables
```


# Save deidentified data {#save}

* Save one copy in feather format for me.

* Save a second copy in CSV format for NACJD.

```{r}
feather::write_feather(
  medstar_compliance_deidentified, 
  "../data/medstar_compliance_deidentified.feather"
)

readr::write_csv(
  medstar_compliance_deidentified,
  "../data/medstar_compliance_deidentified.csv"
)
```

[top](#top)


&nbsp;

-------------------------------------------------------------------------------

```{r echo=FALSE}
sessionInfo()
```
