---
title: "Deidentify Merged MedStar and APS Data"
date: "Created: 2018-10-16 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    css: custom-css.css
---

# Table of contents

* [Overview](#overview)   
* [Load packages and data](#load)   
* [Create an anonymous unique person id number](#person-id)   
* [Deidentify the response numbers for the MedStar legal compliance department](#deidentify-1)   
* [Managing dates](#managing-dates)   
* [Deidentify the merged dataset](#deidentify-2)    
* [Save deidentified version of merged dataset](#save)   
* [Create new study id](#new-id)   
* [Count rows, unique responses, and unique people](#counts)   


# Overview

* Remove identifiers from the merged MedStar/APS data so that it can be uploaded to [NACJD]( https://www.icpsr.umich.edu/icpsrweb/content/NACJD/index.html) - including a unique person id number.


# Load packages and data {#load}

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(bfuncs)
```

Merged MedStar and APS data. Created in data_medstar_aps_merged_01_merge.Rmd

```{r}
medstar_aps_merged <- feather::read_feather("/Volumes/Detect/medstar_aps_merged.feather")
```

```{r}
about_data(medstar_aps_merged) # 101,450 observations and 64 variables
```


## Response ID's from MedStar compliance data

These are the response id numbers that correspond to 911 responses where a report was made to APS during the DETECT pilot phase and MedStar compliance was made aware of the report. 

Data from data_medstar_compliance_01_import.Rmd

```{r}
response_ids <- feather::read_feather("/Volumes/Detect/response_ids.feather")
```

```{r}
about_data(response_ids) # 14 observations and 1 variable
```










# Create an anonymous unique person id number {#person-id}

Some people were treated by MedStar and/or investigated by APS more than one time. Therefore, neither incident pcr number nor case number uniquely identifies people. For many measures, this doesn't matter because we are interested in outcomes at the incident level. However, it does matter for a couple of demographic measures like age, race, etc.

Initially, we attempted to uniquely identify people based on a combination of name and DOB. However, creating a unique identifier this way is complicated by data entry errors (i.e., misspelled names and mistyped dates of birth). For example, john_smith_1936_10_03 and jon_smith_1936_10_03 may be the same person, but would generate two different unique id numbers. 

Instead, we can use fastLink to search for names and DOB's that are likely to be the same person and assign them a unique person id.

If we do this on the entire dataset at once, it is a costly process. It becomes much more tractable if we temporarily reduce the data to one row for each combination of values that we want to investigate. In this case, we want to investigate all unique combinations of name, DOB, and address.

## Setup

```{r}
medstar_aps_merged <- medstar_aps_merged %>% 
  
  # Coerce selected variables to numeric
  mutate_at(
    vars(age, birth_year, birth_mnth, birth_day, address_num, zip),
    as.numeric
  ) %>% 
  
  # Add group variable (combination of values of interest)
  mutate(
    group = paste(first_name, last_name, birth_year, birth_mnth, birth_day, address_street, sep = "_")
  )
```

```{r}
about_data(medstar_aps_merged) # 101,450 observations and 65 variables
```

```{r}
vars_for_matching <- medstar_aps_merged %>% 
  select(first_name, last_name, starts_with("birth"), address_street) %>% 
  names()
```

```{r}
vars_to_exclude <- medstar_aps_merged %>% 
  select(-vars_for_matching) %>% 
  names()
```


## Keep screened only for matching

```{r}
df_screened_only <- medstar_aps_merged %>% filter(detect_data == 1)
```

```{r}
about_data(df_screened_only) # 64,981 observations and 65 variables
```

```{r}
temp <- df_screened_only %>% 
  # Keep one row for every combination of first name, last name, dob, and address
  group_by(group) %>% 
  filter(row_number() == 1) %>% 
  ungroup()
```


# Find matches

```{r}
rpairs_jar <- RecordLinkage::compare.linkage(
  dataset1 = temp, 
  dataset2 = temp,
  strcmp  = c("first_name", "last_name", "address_street"),
  exclude = vars_to_exclude
)
```

```{r}
rpairs_epiwt <- RecordLinkage::epiWeights(rpairs_jar)
```

```{r}
weighted_pairs <- RecordLinkage::getPairs(rpairs_epiwt, max.weight = Inf, min.weight = 0.5)
```

```{r}
weighted_pairs <- weighted_pairs %>% 
  filter(id != "") %>% 
  mutate(
    row = id %>% as.character() %>% as.integer(),
    pair_num = rep(seq(nrow(.) / 2), each = 2),
    Weight = if_else(
      Weight == "", 
      lead(Weight), 
      Weight) %>% 
      as.character() %>% 
      as.numeric() %>% 
      round(6)
  ) %>% 
  select(Weight, row, pair_num, everything(), -id)
```


## Find cut point

Results hidden to protect privacy

```{r eval=FALSE}
weighted_pairs %>% 
  filter(Weight < 0.8) %>% 
  select(Weight, row, pair_num, ends_with("name"), starts_with("birth"), starts_with("address"))
```

```{r}
max_weight <- 0.7693280 - .0001 # Because threshold below is not inclusive
```

```{r}
weighted_pairs <- weighted_pairs %>% filter(Weight >= max_weight)
```

```{r}
check <- weighted_pairs %>%
  select(row) %>% 
  mutate(matches = lead(row)) %>%    # Display horizontally instead of vertically
  filter(row_number() %% 2 == 1) %>% # Drop every other row
  arrange(row) %>% 
  group_by(row) %>% 
  mutate(matches = list(matches)) %>% # Nest all matches for each row
  ungroup() %>% 
  mutate(matches = purrr::map_chr(
    matches, 
    function(x) {
      x = sort(x) # 1,2,3 is the same as 3,2,1
      x = paste(x, collapse = ",") # Convert list to character string
      x
    })
  ) %>% 
  select(matches) %>% 
  distinct() %>% # Reduce to 1 row per group of matches
  mutate(
    person_id = row_number(),
    row = purrr::map( # Turn back into list
      matches, 
      ~ scan(text = ., what = 0L, sep = ",", quiet = TRUE)
    ) 
  ) %>% 
  unnest() %>% 
  select(person_id, row) %>% 
  right_join(
    temp %>% 
      mutate(row = row_number()),
    by = "row"
  )
```

## Data check

How many rows and how many unique ids?

```{r}
nrow(check) # 1,142
```

```{r}
length(unique(check$person_id)) # 1,126
```

So, there are instances when two different rows (by definition that have different values for first name, last, DOB, and/or address because we used distinct above) have the same person id.

Let's view those and make sure the look appropriate (results hidden to protect privacy)

```{r eval=FALSE}
check %>% 
  group_by(person_id) %>% 
  filter(max(row_number()) > 1) %>% 
  arrange(person_id)
```

Those all look appropriate

Now let's look for cases where people weren't matched, but should have been (results hidden to protect privacy) 

```{r eval=FALSE}
# Same name, but different id
check %>%
  mutate(check = paste(first_name, last_name, sep = "_")) %>% 
  group_by(check) %>% 
  filter(max(row_number()) > 1) %>% 
  arrange(check)
```

So, when you look through those there are some good examples of people with the same name being correctly classified as matches, and people with the same name being correctly classified as non-matches. As long, as there isn't missing data, this appears to work really well.


## Adding person id to the data

```{r}
df_screened_only <- check %>%
  select(person_id, group) %>% 
  right_join(df_screened_only, by = "group") %>% 
  select(-group)
```

```{r}
length(unique(df_screened_only$person_id)) # 1,126 matches above
```

```{r}
about_data(df_screened_only) # 64,981 observations and 65 variables
```


## Add non-screened back in

```{r}
medstar_aps_merged <- df_screened_only %>%
  bind_rows(medstar_aps_merged %>% filter(detect_data == 0)) %>% 
  select(-group)
```

At this point, a unique person id (that I feel confident about) exists for every pcr where a screening was conducted.

Rows in the data that aren't associated with a pcr where a screening was conducted don't have a person id. Unfortunately, there isn't anything we can do about it. Without DOB and/or address in these rows, we can't reliably identify unique people. 

The good news is that our interest is really in relationships at the event level as opposed to the person level. For the most part, our inability to accurately identify unique people in the non-DETECT data shouldn't matter much.

## Count of unique people

```{r}
medstar_aps_merged %>% 
  summarise(
    `Unique People` = unique(person_id) %>% 
      length() %>% 
      format(big.mark = ",")
  ) # 1,127
```


## Clean up

```{r}
rm(check, df_screened_only, rpairs_epiwt, rpairs_jar, temp, weighted_pairs, max_weight,
   vars_for_matching, vars_to_exclude)
```

[top](#top)










# Deidentify the response numbers from the MedStar legal compliance department {#deidentify-1}

This has to be done before deidentifying the incient call number.

Check to see if the expectd response numbers from the MedStar compliance data exist in the merged data. 

In data_medstar_detect_screenings_01_import.Rmd there were 8 response id's from the compliance data that had a match in the DETECT screening tool data.

```{r}
about_data(response_ids) # 14 observations and 1 variables
```

anti_join returns all rows in MedStar's legal compliance data that do not have a match in the MedStar DETECT screening items data that was matched to APS cases (results hidden to protect participant privacy).

```{r eval=FALSE}
response_ids %>% 
  anti_join(medstar_aps_merged %>% 
              group_by(incident_pcr_number) %>% 
              filter(row_number() == 1), 
            by = c("response_num" = "incident_call_number")) %>%  
  unique() # 5 incident pcr numbers
```

There are 5 responses id's in the MedStar compliance data that don't appear in the merged MedStar/APS data (results hidden to protect participant privacy). 

I have an email from MedStar from 2016-10-10 outlining the discrepancy. One response id had all nulls for the DETECT screening items (...147). One response id could not be found in the query. The rest were ineligible for the DETECT screening tool because of age.

At this point, there should be 9 response id's from the compliance data that have a match in the merged MedStar/APS data (results hidden to protect participant privacy).

```{r eval=FALSE}
response_ids %>% 
  semi_join(medstar_aps_merged %>% 
              group_by(incident_pcr_number) %>% 
              filter(row_number() == 1), 
            by = c("response_num" = "incident_call_number")) %>% 
  unique() # 9 incident pcr numbers
```

Now we need a way to track these cases going forward without using a number that has meaning in MedStar's database.

We will just add a new variable to the merged data that indicates it was match with the compliance data. 

```{r}
in_compliance_data <- response_ids %>% 
  semi_join(medstar_aps_merged %>% 
              group_by(incident_pcr_number) %>% 
              filter(row_number() == 1), 
            by = c("response_num" = "incident_call_number")) %>% 
  unique() %>% 
  mutate(in_compliance_data = 1L)
```

```{r}
medstar_aps_merged <- medstar_aps_merged %>% 
  left_join(in_compliance_data, by = c("incident_call_number" = "response_num")) %>% 
  mutate(in_compliance_data = if_else(is.na(in_compliance_data), 0L, in_compliance_data))
```

```{r}
about_data(medstar_aps_merged) # 101,450 observations and 66 variables
```


## Data check

At this point, there should be 9 response id's from the compliance data that have a match in the merged MedStar/APS data (results hidden to protect participant privacy).

```{r eval=FALSE}
medstar_aps_merged %>% 
  filter(in_compliance_data == 1) %>% 
  group_by(incident_call_number) %>% 
  filter(row_number() == 1) # 8 rows
```


## Clean up

```{r}
rm(in_compliance_data, response_ids)
```

[top](#top)










# Managing dates {#managing-dates}

## Drop rows from 2015-10-27

There were no screenings done after 2015-10-26. Response data from 2015-10-27 needs to be filtered out.

```{r}
medstar_aps_merged <- medstar_aps_merged %>%
  filter(
    # Keep if it's MedStar incident from 2015-10-26 and earlier
    (!is.na(incident_pcr_number) & as.Date(date_entered) < "2015-10-27") |  
    # OR if it's an APS investigation that wasn't matched with a MedStar record
    (is.na(incident_pcr_number) & is.na(date_entered))
  )
```

```{r}
about_data(medstar_aps_merged) # 98,968 observations and 66 variables
```


## Create validation rule variable

Create a dummy variable that indicates if the validation rule was implemented or not. It has to be done here because we erase the days from dates below.

```{r}
medstar_aps_merged <- medstar_aps_merged %>%
  mutate(validation = if_else(as.Date(date_entered) < "2015-09-28", 0, 1))
```

```{r}
about_data(medstar_aps_merged) # 98,968 observations and 67 variables
```


## Create dataset for figure

Create dataset just to be used for the qualified responses and completed screenings figure.

```{r}
fig_responses_screenings_validation <- medstar_aps_merged %>% 
  filter(!is.na(incident_pcr_number)) %>% 
  group_by(incident_pcr_number) %>% 
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  select(date_entered, detect_data, validation)
```

```{r}
about_data(fig_responses_screenings_validation) # 1,967 observations and 3 variables
```

```{r}
write_csv(fig_responses_screenings_validation, 
          "/Volumes/Detect/fig_responses_screenings_validation.csv")
```


## Remove the day from all remaining date variables

As part of the process of protecting participant anonymity, we will remove the day from all dates.

```{r}
medstar_aps_merged <- medstar_aps_merged %>% 
  mutate(
    date_entered = format(date_entered, "%Y-%m"),
    intake_start = format(intake_start, "%Y-%m")
  )
```

```{r}
about_data(medstar_aps_merged) # 98,968 observations and 67 variables
```


## Clean up

```{r}
rm(fig_responses_screenings_validation)
```

[top](#top)










# Deidentify the merged dataset {#deidentify-2}

We need to create deidentified datasets for [NACJD](https://www.icpsr.umich.edu/icpsrweb/content/NACJD/index.html).


## Save current order of variables

```{r}
current_order <- names(medstar_aps_merged)
```


## Incident call numbers

```{r}
medstar_aps_merged <- medstar_aps_merged %>% 
  select(incident_call_number) %>% 
  unique() %>% 
  mutate(new = row_number()) %>% 
  right_join(medstar_aps_merged, by = "incident_call_number") %>% 
  # Prevent assigning number to NA
  mutate(new = if_else(is.na(incident_call_number), NA_integer_, new)) %>% 
  mutate(incident_call_number = new) %>% 
  select(-new)
```


## Incident pcr numbers

```{r}
medstar_aps_merged <- medstar_aps_merged %>% 
  select(incident_pcr_number) %>% 
  unique() %>% 
  mutate(new = row_number()) %>% 
  right_join(medstar_aps_merged, by = "incident_pcr_number") %>% 
  # Prevent assigning number to NA
  mutate(new = if_else(is.na(incident_pcr_number), NA_integer_, new)) %>%
  mutate(incident_pcr_number = new) %>% 
  select(-new)
```


## Case numbers

```{r}
medstar_aps_merged <- medstar_aps_merged %>%
  select(case_num) %>% 
  unique() %>% 
  mutate(new = row_number()) %>% 
  right_join(medstar_aps_merged, by = "case_num") %>% 
  # Prevent assigning number to NA
  mutate(new = if_else(is.na(case_num), NA_integer_, new)) %>%
  mutate(case_num = new) %>% 
  select(-new) %>% 
  # To force case numbers to start at 1
  mutate(case_num = case_num - 1)
```


## Crew member id

```{r}
medstar_aps_merged <- medstar_aps_merged %>%
  select(crew_member_id) %>% 
  unique() %>% 
  mutate(new = row_number()) %>% 
  right_join(medstar_aps_merged, by = "crew_member_id") %>% 
  # Prevent assigning number to NA
  mutate(new = if_else(is.na(crew_member_id), NA_integer_, new)) %>%
  mutate(crew_member_id = new) %>% 
  select(-new)
```


## Perp id

```{r}
medstar_aps_merged <- medstar_aps_merged %>%
  select(perp_id) %>% 
  unique() %>% 
  mutate(new = row_number()) %>% 
  right_join(medstar_aps_merged, by = "perp_id") %>% 
  # Prevent assigning number to NA
  mutate(new = if_else(is.na(perp_id), NA_integer_, new)) %>%
  mutate(perp_id = new) %>% 
  select(-new) %>% 
  # To force perp ids to start at 1
  mutate(perp_id = perp_id - 1)
```


## Remove remaining identifiers

```{r}
medstar_aps_merged <- medstar_aps_merged %>% 
  select(-first_name, -last_name, -birth_year, -birth_mnth, -birth_day, -address_num, 
         -address_street, -city, -zip, -first_name_02, -response_date, -aps_report_num,
         -full_name, -address, -state, -dob)
```

```{r}
about_data(medstar_aps_merged) # 98,968 observations and 51 variables
```


## Reorder variables

```{r}
current_order <- setdiff(
  current_order, 
  c("first_name", "last_name", "birth_year", "birth_mnth", "birth_day",
    "address_num", "address_street", "city", "zip", "first_name_02", 
    "response_date", "aps_report_num", "full_name", "address", "state", "dob")
)
```


```{r}
medstar_aps_merged <- medstar_aps_merged %>%
  select(current_order)
```

```{r}
about_data(medstar_aps_merged) # 98,968 observations and 51 variables
```


## Clean up

```{r}
rm(current_order)
```

[top](#top)










# Create new study id {#new-id}

In the merged data there are incidents where people were treated by MedStar -- producing an incident pcr number -- but were not investigated by APS.

Additionally, there are investigations that were conducted by APS -- producing a case number -- but were not screened by MedStar.

Therefore, neither incident pcr number nor case number can be used to subset the data to one row per incident/case.

I want to create a new id that can be used for that purpose.

## Data check

```{r}
medstar_aps_merged <- medstar_aps_merged %>% 
  mutate(pcr_case = paste(incident_pcr_number, case_num, sep = "_"))
```

```{r}
medstar_aps_merged %>% 
  filter(pcr_case == "NA_NA")
```

Every analysis unit has either an incident pcr number or a case number.

## Create study id

```{r}
medstar_aps_merged <- medstar_aps_merged %>% 
  left_join(
    medstar_aps_merged %>% 
      group_by(pcr_case) %>% 
      filter(row_number() == 1) %>% 
      ungroup() %>% 
      mutate(id = row_number()) %>% 
      select(pcr_case, id),
    by = "pcr_case"
  ) %>% 
  select(id, everything()) %>% 
  select(-pcr_case)
```

```{r}
about_data(medstar_aps_merged) # 98,968 observations and 52 variables
```


[top](#top) 










# Save deidentified version of merged dataset {#save}

```{r}
feather::write_feather(
  medstar_aps_merged, 
  "../data/detect_5wk_1.feather"
)
```

```{r}
readr::write_csv(
  medstar_aps_merged, 
  "../data/detect_5wk_1.csv"
)
```

[top](#top)










# Count rows, unique responses, and unique people {#counts}

Not including missing

```{r}
medstar_aps_merged %>% 
  summarise(
    `Total Rows` = n(), # 98,968
    `Study IDs` = id %>% unique() %>% is.na() %>% `!` %>% sum(), # 2,611
    `Unique Incidents` = incident_call_number %>% unique() %>% is.na() %>% `!` %>% sum(), # 1,964
    `Unique PCR` = incident_pcr_number %>% unique() %>% is.na() %>% `!` %>% sum(), # 1,967
    `Unique Case Numbers` = case_num %>% unique() %>% is.na() %>% `!` %>% sum(), # 679
    `Unique People` = person_id %>% unique() %>% is.na() %>% `!` %>% sum() # 1,126
  ) %>% 
  mutate_all(format, big.mark = ",")
```

[top](#top)


## Clean up

```{r}
rm(list = ls())
```

&nbsp;

-------------------------------------------------------------------------------

```{r echo=FALSE}
sessionInfo()
```
