---
title: "Preprocess merged data for analysis"
date: "Created: 2018-10-27 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    css: custom-css.css
---

# Table of contents

* [Overview](#overview) 
* [Load packages and data](#load)   
* [Variable management](#var-manage)    
* [Save preprocessed data](#save)    

# Overview

In this file we take the MedStar DETECT and APS investigations data that we previously merged together and do a little bit of preprocessing for later analysis.


# Load packages and data {#load}

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(bfuncs)
```

```{r}
detect_5wk_1 <- feather::read_feather("../data/detect_5wk_1.feather")
```

```{r}
about_data(detect_5wk_1) # 101,450 observations and 51 variables
```










# Variable management {#var-manage}

Because it is possible for there to be more than one allegation associated with each APS investigation, some investigations have more than one row in the data. It is also possible to have more than one instance of the DETECT screening tool used (incident pcr number) per APS investigation (case number).

We want a single outcome per investigation/pcr combination in order to create a confusion matrix for each screening item.

To do that, we will create a new variable called "any_valid" that is equal to 1 if the disposition of _any_ allegation was valid, and 0 if the disposition of _all_ allegations were invalid. 

## Create any_valid

Create new outcome variable (any_valid) that is equal to 1 when any type of abuse is validated for a given response pair (DETECT screen / APS investigation match). For response pairs with multiple dispositions:   
  + If even one disposition is valid, then any valid equals 1   
  + If all dispositions are invalid, then any valid equals 0   
  + If all dispositions are other, then any valid equals NA
  + If one disposition is other, and the rest are invalid, then any valid equals NA
  
```{r}
detect_5wk_1 <- detect_5wk_1 %>%
  group_by(id) %>% 
  mutate(
    any_valid = case_when(
      any(disposition == "Valid")   ~ 1L,
      all(disposition == "Invalid") ~ 0L,
      TRUE                          ~ NA_integer_
    ) 
  ) %>% 
  ungroup()
```

## Data check

```{r}
detect_5wk_1 %>% 
  select(id, incident_pcr_number, case_num, disposition, any_valid) %>% 
  distinct()
```

```{r}
about_data(detect_5wk_1) # 101,450 observations and 52 variables
```


## Recode DETECT responses to NA

* We will recode "Not applicable or No Caregiver" to NA. The only variable that contained this response was no_talk51. This response was given 3 times.

* We will also recode "Don’t Know" to NA.

```{r}
detect_5wk_1 <- detect_5wk_1 %>% 
  map_at(
    .x  = .,
    .at = select(detect_5wk_1, unusual_odor36:adls61) %>% names(),
    .f  = fct_recode, NULL = "Not applicable or No Caregiver"
  ) %>% 
  map_at(
    .x  = .,
    .at = select(detect_5wk_1, unusual_odor36:adls61) %>% names(),
    .f  = fct_recode, NULL = "Don’t Know"
  ) %>% 
  as_tibble()
```

```{r}
about_data(detect_5wk_1) # 101,450 observations and 52 variables
```

[top](#top)










# Save preprocessed data {#save}

```{r}
feather::write_feather(detect_5wk_1, "../data/detect_5wk_2.feather")
```

[top](#top)

&nbsp;

-------------------------------------------------------------------------------

```{r echo=FALSE}
sessionInfo()
```
