---
title: "Screening and Reporting Flow Diagram"
date: "Created: 2018-11-01 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    css: custom-css.css
---

# Table of contents

* [Overview](#overview)   
* [Calculate numbers of screenings and reports](#calculations)   
* [Create flow chart](#create)   


# Overview

Create a flow chart that graphically illustrates the steps in the screening and reporting process during the 5-week pilot study.

Doing this as a Notebook right now. We can change it to another output format later if we need to.

Helpful website: https://www.stat.auckland.ac.nz/~paul/R/Diagram/diagram.pdf


# Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
library(bfuncs)
library(DiagrammeR)
library(grid)
library(Gmisc)
```


## Deidentified 5-week pilot data

Created in data_medstar_aps_merged_03_preprocess_for_analysis.Rmd

```{r}
detect_5wk <- feather::read_feather("../data/detect_5wk.feather")
```

```{r}
about_data(detect_5wk) # 98,968 observations and 56 variables
```










# Calculate numbers of screenings and reports {#calculations}

Calculated measures to include in the diagram boxes along with descriptive text.

## Reduce data to 1 row per pcr

```{r}
detect_1id <- detect_5wk %>% 
  group_by(id) %>% 
  filter(row_number() == 1) %>% 
  ungroup()
```

```{r}
about_data(detect_1id) # 2,611 observations and 56 variables
```


## Eligibility

Number of MedStar incident PCRs for a person who was 65 years old or older, treated in their residence, and treated during the 5-week DETECT pilot study (2015-09-17 -- 2015-10-26).

Only rows that correspond to a MedStar response (as opposed to a APS investigation that wasn't paired with a response) should be count as eligible.

```{r}
eligible   <- detect_1id %>% filter(!is.na(incident_pcr_number))
n_eligible <- nrow(eligible) %>% format(big.mark = ",") # 1,967
```

```{r}
ineligible   <- detect_1id %>% filter(is.na(incident_pcr_number))
n_ineligible <- nrow(ineligible) %>% format(big.mark = ",") # 644
```

```{r}
rm(ineligible)
```


## Screened

At how many of the eligible MedStar responses was 1 or more items from the DETECT tool used to screen for EA?

```{r}
screened   <- eligible %>% filter(detect_data == 1)
n_screened <- nrow(screened) %>% format(big.mark = ",") # 1,247
```

```{r}
not_screened   <- eligible %>% filter(detect_data == 0)
n_not_screened <- nrow(not_screened) %>% format(big.mark = ",") # 720
```

```{r}
rm(eligible, not_screened)
```


## Screen positive

How many times was one or more DETECT screening items positive?

```{r}
screen_pos   <- screened %>% filter(yes_any == 1)
n_screen_pos <- nrow(screen_pos) %>% format(big.mark = ",") # 209
```

```{r}
screen_neg   <- screened %>% filter(yes_any == 0)
n_screen_neg <- nrow(screen_neg) %>% format(big.mark = ",") # 1,038
```

```{r}
rm(screened, screen_neg)
```


## Report to APS

How many positive screens were investigated by APS?

```{r}
report_aps   <- screen_pos %>% filter(!is.na(case_num))
n_report_aps <- nrow(report_aps) %>% format(big.mark = ",") # 20
```

```{r}
no_report_aps   <- screen_pos %>% filter(is.na(case_num))
n_no_report_aps <- nrow(no_report_aps) %>% format(big.mark = ",") # 189
```

```{r}
rm(screen_pos, no_report_aps)
```


## Report to MedStar compliance

How many reports to APS was MedStar compliance made aware of?

```{r}
report_compliance   <- report_aps %>% filter(in_compliance_data == 1)
n_report_compliance <- nrow(report_compliance) %>% format(big.mark = ",") # 7
```

Why aren't their 9?

```{r eval=FALSE}
detect_1id %>% filter(in_compliance_data == 1)
```

Two of the incidents (id 491 and id 2608 weren't matched to APS investigations).

```{r}
no_report_compliances   <- report_aps %>% filter(in_compliance_data == 0)
n_no_report_compliances <- nrow(no_report_compliances) %>% format(big.mark = ",") # 13
```

```{r}
rm(report_aps, report_compliance, no_report_compliances)
```

[top](#top)










# Create flow chart {#create}

```{r fig.height=4, fig.width=5}
# =============================================================================
# Open graphical device (the type output you want)
# =============================================================================
jpeg(
  '../images/figure_screening_and_reporting_flow_diagram.jpg',
  height = 672,
  width = 720
)

# =============================================================================
# Create Viewport
# =============================================================================
grid.newpage()
chart_width <- grid::unit(7.5, "inch")
chart_height <- grid::unit(7, "inch")
pushViewport(viewport(width = chart_width, height = chart_height))

# =============================================================================
# Calculations for box spacing
# =============================================================================
splits <- 4
n_vert_layers <- 5
n_horz_layers <- (splits * 2) + 1
vert_breaks <- 1/n_vert_layers
horz_breaks <- 1/n_horz_layers
y_coords <- seq(vert_breaks * n_vert_layers, vert_breaks, -vert_breaks)
x_coords <- seq(horz_breaks, horz_breaks * n_horz_layers, horz_breaks)

# =============================================================================
# Create text boxes and print to viewport
# =============================================================================

(eligible <- boxGrob(
  label = paste("Eligible\n N =", n_eligible),
  x = x_coords[8],
  y = y_coords[1],
  box_gp = gpar(fill = "white")
))

(screened <- boxGrob(
  label = paste("Screened\n N =", n_screened),
  x = x_coords[6],
  y = y_coords[2],
  box_gp = gpar(fill = "white")
))

(not_screened <- boxGrob(
  label = paste("Not Screened\n N =", n_not_screened),
  x = x_coords[9],
  y = y_coords[2],
  box_gp = gpar(fill = "white")
))

(screen_pos <- boxGrob(
  label = paste("Screened Positive\n N =", n_screen_pos),
  x = x_coords[4],
  y = y_coords[3],
  box_gp = gpar(fill = "white")
))

(screen_neg <- boxGrob(
  label = paste("Screened Negative\n N =", n_screen_neg),
  x = x_coords[7],
  y = y_coords[3],
  box_gp = gpar(fill = "white")
))

(report <- boxGrob(
  label = paste("Report to APS\n N =", n_report_aps),
  x = x_coords[2],
  y = y_coords[4],
  box_gp = gpar(fill = "white")
))

(no_report <- boxGrob(
  label = paste("No Report to APS\n N =", n_no_report_aps),
  x = x_coords[5],
  y = y_coords[4],
  box_gp = gpar(fill = "white")
))

(compliance <- boxGrob(
  label = paste("Inform MedStar Compliance\n N =", n_report_compliance),
  x = x_coords[1] - 0.1,
  y = y_coords[5],
  box_gp = gpar(fill = "white")
))

(no_compliance <- boxGrob(
  label = paste("Did Not Inform MedStar Compliance\n N =", n_no_report_compliances),
  x = x_coords[3] + 0.1,
  y = y_coords[5],
  box_gp = gpar(fill = "white")
))

# Text elements for the timeline
# start_date <- grid.text(label = "September 17, 2015", x = 0, y = 0.05, hjust = 0.75)
# end_date   <- grid.text(label = "October 26, 2015", x = 1, y = 0.05)

# =============================================================================
# Connect text boxes
# =============================================================================

# First, create a reusable set of parameters for the arrow
arrow_parms <- grid::arrow(angle = 30, length = unit(4, "mm"), type = "closed")

# Start building connections
connectGrob(eligible, screened, type = "N", arrow_obj = arrow_parms)
connectGrob(eligible, not_screened, type = "N", arrow_obj = arrow_parms)
connectGrob(screened, screen_pos, type = "N", arrow_obj = arrow_parms)
connectGrob(screened, screen_neg, type = "N", arrow_obj = arrow_parms)
connectGrob(screen_pos, report, type = "N", arrow_obj = arrow_parms)
connectGrob(screen_pos, no_report, type = "N", arrow_obj = arrow_parms)
connectGrob(report, compliance, type = "N", arrow_obj = arrow_parms)
connectGrob(report, no_compliance, type = "N", arrow_obj = arrow_parms)

# Create timeline
# grid.lines(
#   x = c(0.06, 0.9), 
#   y = c(0.05, 0.05), 
#   arrow = arrow(
#     angle = 30, 
#     length = unit(4, "mm"), 
#     ends = "both", 
#     type = "closed"
#   ),
#   gp = gpar(fill = "black")
# )

popViewport()

# =============================================================================
# Stop writing to png file
# =============================================================================
dev.off()
```

[top](#top)


&nbsp;

-------------------------------------------------------------------------------

```{r echo=FALSE}
sessionInfo()
```
